PLAE enables quick and easy searching across one million ocular transcriptomes

### 2022-03-30 manuscript_00.Rmd

## Pan mouse and human non-ocular references

The Tabula Muris and

@@

While scEiaD does not have spatial scRNA seq datasets as of 2022, we do leverage existing knowledge of the structure of the retina to show *in silico* visualization of genes by cell type / layer in a cartoon cross section of the retina. The structure of the cartoon retina was based on the human retina. This unique visualization displays the major cell types of the retina and the RPE / choroid layer behind it. The cell types are colored by the relative amounts of expression in the user selected gene. Like all visualizations in PLAE, the user can filter to only display data from certain metadata characteristics, like species or publication.

@@

## Pan-study fovea vs peripheral photoreceptor identifies N genes consistently differentially expressed

>>>>

## Differential gene testing

Pseudobulk matrices were created for each organism with genes as the rows and study_accession - [test] as the columns; [test] is either cluster, CellType, or CellType_predict with the R package scater function XXX. Genes were removed which had zero expression across all cells. Columns were not retained if they were composed of fewer than 20 total cells. These pseudobulk matrices were used as the input for DESeq2 differential testing. The design of the test was "~study_accession + [test]" where, again, test is either cluster, CellType, or CellType_predict. This design sets the study_accession as the covariate, thus removing effects that are specific to a small number of studies. As the differential testing was done for each individual species, we aligned the separate result tables by the Ensembl gene ID to create a unified table. Values (e.g. pvalue) specific

@@

While scEiaD does not have spatial scRNA seq datasets as of 2022, we do leverage existing knowledge of the structure of the retina to show *in silico* visualization of genes by cell type / layer in a cartoon cross section of the retina. The structure of the cartoon retina was based on the human retina. This unique visualization displays the major cell types of the retina and the RPE / choroid layer behind it. The cell types are colored by the relative amounts of expression in the user selected gene. Like all visualizations in PLAE, the user can use powerful filtering options to only display data from certain metadata characteristics, like species or publication.

@@

To aid in in the analysis of ocular gene expression patterns we provide pre-computed differential tests run 
## Pan-study fovea vs peripheral photoreceptor identifies N genes consistently differentially expressed
